---
title: "Project Part 2"
output: 
  html_notebook: 
    theme: cerulean
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

***

This notebook contains the code samples found in Chapter 3, Section 5 of [Deep Learning with R](https://www.manning.com/books/deep-learning-with-r). Note that the original text features far more content, in particular further explanations and figures: in this notebook, you will only find source code and related comments.

***

# Data Exploration & Preparation 
* Our goal in the second part of the assignment is to predict how good a (new) customer will pay 
back their credit card depts. In the data set application data from current customers (the first 18 
attributes) together with their status (last attribute; target) are given.  
* The attributes from the applications are 

Attribute Name | Explanation | Remarks
------------- | ------------- | -------------
ID | Client | number 
CODE_GENDER | Gender | 
FLAG_OWN_CAR | Is there a car | 
FLAG_OWN_REALTY | Is there a property | 
CNT_CHILDREN | Number of children | 
AMT_INCOME_TOTAL | Annual income | 
NAME_INCOME_TYPE | Income category | 
NAME_EDUCATION_TYPE | Education level | 
NAME_FAMILY_STATUS | Marital status | 
NAME_HOUSING_TYPE | Way of living | 
DAYS_BIRTH | Birthday | Count backwards from current day (0), -1 means yesterday 
DAYS_EMPLOYED | Start date of employment | Count backwards from current day(0). If positive, it means the person unemployed. 
FLAG_MOBIL | Is there a mobile phone | 
FLAG_WORK_PHONE | Is there a work phone | 
FLAG_PHONE | Is there a phone | 
FLAG_EMAIL | Is there an email | 
OCCUPATION_TYPE | Occupation | 
CNT_FAM_MEMBERS | Family size | 

* The last attribute status contains the “pay-back behavior”, i.e. when did that customer pay back 
their depts: 
  + 0: 1-29 days past due 
  + 1: 30-59 days past due 
  + 2: 60-89 days overdue 
  + 3: 90-119 days overdue 
  + 4: 120-149 days overdue 
  + 5: Overdue or bad debts, write-offs for more than 150 days 
  + C: paid off that month 
  + X: No loan for the month 
Please note: We are learning only the pay-back behavior. The decision, i.e. if we accept a customer or 
not, is done in another process step – not here!  

**Single-label, Multiclass classification in page 73 in [Deep Learning with R](https://www.manning.com/books/deep-learning-with-r)**

***

# Main task 
* Design your network. Why did you use a feed-forward network, or a convolutional or recursive 
network – and why not?  
* Use k-fold validation (with k = 10) to find the best hyperparameters for your network. 
* Use the average of the accuracy to evaluate the performance of your trained network. 
* Find a “reasonable” good model. Argue why that model is reasonable. If you are not able to find a 
reasonable good model, explain what you all did to find a good model and argue why you think 
that’s not a good model.  
* Save your trained neural network with save_model_hdf5. Also save your data sets you used 
for training, testing and validation. 

***

# Some hints 
* Data preprocessing is easier here; no feature engineering is needed. 
* You may be able to reuse parts of the exercises we used in our examples during lectures. 
* All in- and output values need to be floating numbers (or integers in exceptions) in the range of 
[0,1]. 
* Please note that a neural network expects a R matrix or vector, not data frames. Transform your 
data (e.g. a data frame) into a matrix with data.matrix if needed.  
* There are some models which show an accuracy higher than 90% (!) for training (and test) data – 
after learning more than 1000 epochs. 

***

# Important notes
* Spaces must be removed in between '```{r}' and '```', else an error with '<!-- rnb-source-end -->' will be produced

## Data import


```{r}
library(here)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tensorflow)
library(tfdatasets)
library(keras)
library(caret)
#LOAD DATA
setwd(getwd())
dataIn = "../Data/Dataset-part-2.csv"
data_in <- read.csv(dataIn,header = TRUE, sep =',')
View(data_in)
data <- data.frame(data_in)
summary(data)
```
##Cleanup
```{r}
data <- data %>% select(-ID)
cols <- c("CODE_GENDER","FLAG_OWN_CAR","FLAG_OWN_REALTY","NAME_INCOME_TYPE","NAME_EDUCATION_TYPE", "NAME_FAMILY_STATUS", "NAME_HOUSING_TYPE","FLAG_MOBIL","FLAG_WORK_PHONE","FLAG_PHONE","FLAG_EMAIL", "OCCUPATION_TYPE","status")
cols
data[cols] <- lapply(data[cols],factor)

# One-Hot-Encoding for CODE_GENDER
dmy <- dummyVars(" ~ CODE_GENDER", data = data)
trsf <- data.frame(predict(dmy, newdata = data))
data <- cbind(data, trsf)
data <- data %>% select(-CODE_GENDER)

summary(data)
```
##K-Fold-Validation
```{r}
   # 75% of the sample size
  smp_size <- floor(0.75 * nrow(data))
  
  # set the seed to make your partition reproducible
  set.seed(123)
  data_ind <- sample(seq_len(nrow(data)), size = smp_size)
  train <- as_tibble(data[data_ind, ])
  test <- as_tibble(data[-data_ind, ])


  
  
```
```{r}
  ## Take Model1 and Scale data
  ##scale and preprocess training and test data
  library(caret)
  pre_proc_val <- preProcess(train, method = c("center", "scale"))  
  train = predict(pre_proc_val, train)
  test = predict(pre_proc_val, test)
  train <- tensor_slices_dataset(train)
  test <- tensor_slices_dataset(test)
  summary(train)
  
```


```{r, results='hide'}
library(keras)
#https://www.tensorflow.org/tutorials/structured_data/feature_columns

#center and scale --> min max
#one hot encoding
# Matrix and vector

# model = tf.keras.Sequential([
#   feature_layer,
#   layers.Dense(128, activation='relu'),
#   layers.Dense(128, activation='relu'),
#   layers.Dropout(.1),
#   layers.Dense(1)
# ])
# 
# model.compile(optimizer='adam',
#               loss=tf.keras.losses.BinaryCrossentropy(from_logits=True),
#               metrics=['accuracy'])
# 
# model.fit(train_ds,
#           validation_data=val_ds,
#           epochs=10)

```



```{r}
str(train_data[[1]])
```