---
title: "Project Part 2"
output: 
  html_notebook: 
    theme: cerulean
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

***

This notebook contains the code samples found in Chapter 3, Section 5 of [Deep Learning with R](https://www.manning.com/books/deep-learning-with-r). Note that the original text features far more content, in particular further explanations and figures: in this notebook, you will only find source code and related comments.

***

# Data Exploration & Preparation 
* Our goal in the second part of the assignment is to predict how good a (new) customer will pay 
back their credit card depts. In the data set application data from current customers (the first 18 
attributes) together with their status (last attribute; target) are given.  
* The attributes from the applications are 

Attribute Name | Explanation | Remarks
------------- | ------------- | -------------
ID | Client | number 
CODE_GENDER | Gender | 
FLAG_OWN_CAR | Is there a car | 
FLAG_OWN_REALTY | Is there a property | 
CNT_CHILDREN | Number of children | 
AMT_INCOME_TOTAL | Annual income | 
NAME_INCOME_TYPE | Income category | 
NAME_EDUCATION_TYPE | Education level | 
NAME_FAMILY_STATUS | Marital status | 
NAME_HOUSING_TYPE | Way of living | 
DAYS_BIRTH | Birthday | Count backwards from current day (0), -1 means yesterday 
DAYS_EMPLOYED | Start date of employment | Count backwards from current day(0). If positive, it means the person unemployed. 
FLAG_MOBIL | Is there a mobile phone | 
FLAG_WORK_PHONE | Is there a work phone | 
FLAG_PHONE | Is there a phone | 
FLAG_EMAIL | Is there an email | 
OCCUPATION_TYPE | Occupation | 
CNT_FAM_MEMBERS | Family size | 

* The last attribute status contains the “pay-back behavior”, i.e. when did that customer pay back 
their depts: 
  + 0: 1-29 days past due 
  + 1: 30-59 days past due 
  + 2: 60-89 days overdue 
  + 3: 90-119 days overdue 
  + 4: 120-149 days overdue 
  + 5: Overdue or bad debts, write-offs for more than 150 days 
  + C: paid off that month 
  + X: No loan for the month 
Please note: We are learning only the pay-back behavior. The decision, i.e. if we accept a customer or 
not, is done in another process step – not here!  


***

# Main task 
* Design your network. Why did you use a feed-forward network, or a convolutional or recursive 
network – and why not?  
* Use k-fold validation (with k = 10) to find the best hyperparameters for your network. 
* Use the average of the accuracy to evaluate the performance of your trained network. 
* Find a “reasonable” good model. Argue why that model is reasonable. If you are not able to find a 
reasonable good model, explain what you all did to find a good model and argue why you think 
that’s not a good model.  
* Save your trained neural network with save_model_hdf5. Also save your data sets you used 
for training, testing and validation. 

***

# Some hints 
* Data preprocessing is easier here; no feature engineering is needed. 
* You may be able to reuse parts of the exercises we used in our examples during lectures. 
* All in- and output values need to be floating numbers (or integers in exceptions) in the range of 
[0,1]. 
* Please note that a neural network expects a R matrix or vector, not data frames. Transform your 
data (e.g. a data frame) into a matrix with data.matrix if needed.  
* There are some models which show an accuracy higher than 90% (!) for training (and test) data – 
after learning more than 1000 epochs. 

***

# Important notes
* Single-label, Multiclass classification problem on page 73 in [Deep Learning with R](https://www.manning.com/books/deep-learning-with-r)
* Spaces must be removed in between '```{r}' and '```', else an error with '<!-- rnb-source-end -->' will be produced
* K-Fold Validation on page 83ff and 94ff in [Deep Learning with R](https://www.manning.com/books/deep-learning-with-r)
* Page 110, use Last-Layer activation softmax and loss function categorical_crossentropy
* Convolutional network ausgeschlossen, weil hauptsächlich Pattern recognition/image classification
* Recursive ausgeschlossen, weil hauptsächlich für TimeSeries-Vorhersagen verwendet, oder für Vorhersagen
* Feed-Forward, weil Classification-Task

***

## Data import
```{r}
#install.packages("tidymodels")
#install.packages("themis")
library(here)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tensorflow)
library(tfdatasets)
library(tidymodels)
library(keras)
library(caret)
library(themis)
#LOAD DATA
setwd(getwd())
dataIn = "../Data/Dataset-part-2.csv"
data_in <- read.csv(dataIn,header = TRUE, sep =',')
#View(data_in)
data <- data.frame(data_in)
summary(data)
plot(data$status)
```
##Cleanup
```{r}
# Check for duplicates 
sum(duplicated(data))
# No duplicates

#Remove ID (irrelevant) and FLAG_MOBIL (always 1)
data <- data %>% select(-ID, -FLAG_MOBIL)
cols <- c("CODE_GENDER","FLAG_OWN_CAR","FLAG_OWN_REALTY","NAME_INCOME_TYPE","NAME_EDUCATION_TYPE", "NAME_FAMILY_STATUS", "NAME_HOUSING_TYPE","FLAG_WORK_PHONE","FLAG_PHONE","FLAG_EMAIL", "OCCUPATION_TYPE","status")
cols
data[cols] <- lapply(data[cols],factor)

# Replacing empty values with "Unknown"
levels(data$OCCUPATION_TYPE) <- c(levels(data$OCCUPATION_TYPE), "Unknown")
data$OCCUPATION_TYPE[is.na(data$OCCUPATION_TYPE)] <- "Unknown"

# Replacing C and X in Status
levels(data$status)[levels(data$status)=="C"] <- "6"
#data$status[data$status == "X"] <- 7
levels(data$status)[levels(data$status)=="X"] <- "7"
# #Convert factors into numericals
# data %<>% mutate_if(is.factor, as.numeric)

summary(data)
```

# Preprocessing
```{r Create a recipe for preproc}
set.seed(1)
trainIndex <- initial_split(data, prop = 0.7, strata = status) 
trainingSet <- training(trainIndex)
testSet <- testing(trainIndex)
status_folds <- vfold_cv(trainingSet, v = 10, strata = status)
status_folds
```


```{r Create a recipe for preproc2}
set.seed(5)
preprocRecipe <-
  recipe(status ~., data = data) %>%
  step_dummy(all_nominal(), -status,  one_hot = TRUE) %>%
  step_range(all_predictors(), -all_nominal(), min = 0, max = 1)%>%
 # step_downsample(status, over_ratio = 1) %>%
  step_smote(status, over_ratio = 0.5, skip=TRUE) %>%
 # step_smotenc(status, over_ratio = 1) %>%
 #step_adasyn(status, over_ratio = 1) %>%
 #step_nearmiss(status, over_ratio = 1) %>%
   
  step_dummy(status,  one_hot = TRUE)# %>%


```

# In this step the above defined receipt is extracted using the `prep()` function, and then use the `bake()` function to transform a set of data based on that recipe.
```{r Prep and bake the defined recipe}
# retain = TRUE and new_data = NULL ensures that pre-processed trainingSet is returned 
trainingSet_processed <- preprocRecipe %>%
  prep(trainingSet, retain = TRUE) %>%
  bake(new_data = NULL)
testSet_processed <- preprocRecipe %>%
  prep(testSet) %>%
  bake(new_data =testSet)

#summary(trainingSet_processed)
```

#OneHotEncoding
```{r}

# # One-Hot-Encoding for CODE_GENDER
# dmy <- dummyVars(" ~ CODE_GENDER", data = data)
# trsf <- data.frame(predict(dmy, newdata = data)) 
# data <- cbind(data, trsf)
# data <- data %>% select(-CODE_GENDER)
# 
# # One-Hot-Encoding for FLAG_OWN_CAR 
# dmyCar <- dummyVars(" ~ FLAG_OWN_CAR", data = data)
# trsfCar <- data.frame(predict(dmyCar, newdata = data))
# data <- cbind(data, trsfCar)
# data <- data %>% select(-FLAG_OWN_CAR)
# 
# # One-Hot-Encoding for FLAG_OWN_REALTY   
# dmyRealty <- dummyVars(" ~ FLAG_OWN_REALTY", data = data)
# trsfRealty <- data.frame(predict(dmyRealty, newdata = data))
# data <- cbind(data, trsfRealty)
# data <- data %>% select(-FLAG_OWN_REALTY)
# 
# # One-Hot-Encoding for NAME_INCOME_TYPE   
# dmyNIT <- dummyVars(" ~ NAME_INCOME_TYPE", data = data)
# trsfNIT <- data.frame(predict(dmyNIT, newdata = data))
# data <- cbind(data, trsfNIT)
# data <- data %>% select(-NAME_INCOME_TYPE)
# 
# # Factoring NAME_EDUCATION_TYPE as it is ordinal
# # unique(data$NAME_EDUCATION_TYPE)
# # "Secondary / secondary special", "Higher education", "Incomplete higher", "Lower secondary", "Academic degree"
# # Ranking:
# # 'Lower secondary', 'Secondary / secondary special', 'Incomplete higher', 'Higher education', 'Academic degree'
# data$NAME_EDUCATION_TYPE <- as.numeric(factor(data$NAME_EDUCATION_TYPE, order=TRUE, levels=c('Lower secondary', 'Secondary / secondary special', 'Incomplete higher', 'Higher education', 'Academic degree')))
# #dmyNET <- dummyVars(" ~ NAME_EDUCATION_TYPE", data = data)
# #trsfNET <- data.frame(predict(dmyNET, newdata = data))
# #data <- cbind(data, trsfNET)
# #data <- data %>% select(-NAME_EDUCATION_TYPE)
# 
# # One-Hot-Encoding for NAME_FAMILY_STATUS   
# dmyNFS <- dummyVars(" ~ NAME_FAMILY_STATUS", data = data)
# trsfNFS <- data.frame(predict(dmyNFS, newdata = data))
# data <- cbind(data, trsfNFS)
# data <- data %>% select(-NAME_FAMILY_STATUS)
# 
# # One-Hot-Encoding for NAME_HOUSING_TYPE   
# dmyNHT <- dummyVars(" ~ NAME_HOUSING_TYPE", data = data)
# trsfNHT <- data.frame(predict(dmyNHT, newdata = data))
# data <- cbind(data, trsfNHT)
# data <- data %>% select(-NAME_HOUSING_TYPE)
# 
# # Remove FLAG_MOBIL, it is always 1
# #data <- data %>% select(-FLAG_MOBIL)
# 
# # One-Hot-Encoding for FLAG_WORK_PHONE  
# # Not needed, already 1 or 0
# # dmyFWP <- dummyVars(" ~ FLAG_WORK_PHONE", data = data)
# # trsfFWP <- data.frame(predict(dmyFWP, newdata = data))
# # data <- cbind(data, trsfFWP)
# # data <- data %>% select(-FLAG_WORK_PHONE)
# 
# # One-Hot-Encoding for FLAG_PHONE  
# # Not needed, already 1 or 0
# # dmyFP <- dummyVars(" ~ FLAG_PHONE", data = data)
# # trsfFP <- data.frame(predict(dmyFP, newdata = data))
# # data <- cbind(data, trsfFP)
# # data <- data %>% select(-FLAG_PHONE)
# 
# # One-Hot-Encoding for FLAG_EMAIL 
# # Not needed, already 1 or 0
# # dmyFE <- dummyVars(" ~ FLAG_EMAIL", data = data)
# # trsfFE <- data.frame(predict(dmyFE, newdata = data))
# # data <- cbind(data, trsfFE)
# # data <- data %>% select(-FLAG_EMAIL)
# 
# # One-Hot-Encoding for OCCUPATION_TYPE
# #data$OCCUPATION_TYPE
# #levels(data$OCCUPATION_TYPE)
# levels(data$OCCUPATION_TYPE) <- c(levels(data$OCCUPATION_TYPE), "Unknown")
# data$OCCUPATION_TYPE[is.na(data$OCCUPATION_TYPE)] <- "Unknown"
# dmyOT <- dummyVars(" ~ OCCUPATION_TYPE ", data = data)
# trsfOT <- data.frame(predict(dmyOT, newdata = data))
# data <- cbind(data, trsfOT)
# data <- data %>% select(-OCCUPATION_TYPE )
# 
# #summary(data)
# 
# ### Normalizing data
# 
# ## Does not work for single columns
# # preprocessParams <- preProcess(data$CNT_CHILDREN, method=c("center", "scale"))
# # summarize transform parameters
# # print(preprocessParams)
# # transform the dataset using the parameters
# # transformed <- predict(preprocessParams, iris[,1:4])
# 
# ## Does not work for single columns
# # mean <- apply(data, 2, mean)
# # std <- apply(data, 2, sd)
# # dataScaled <- scale(data$CNT_CHILDREN, center = mean, scale = std)
# 
# # ## Based on https://www.learnbymarketing.com/tutorials/neural-networks-in-r-tutorial/
# # #Might not be needed if done in preprocessing 
# # data$CNT_CHILDREN <- (data$CNT_CHILDREN-min(data$CNT_CHILDREN)) / (max(data$CNT_CHILDREN)-min(data$CNT_CHILDREN))
# # data$DAYS_BIRTH <- (data$DAYS_BIRTH-min(data$DAYS_BIRTH)) / (max(data$DAYS_BIRTH)-min(data$DAYS_BIRTH))
# # data$DAYS_EMPLOYED <- (data$DAYS_EMPLOYED-min(data$DAYS_EMPLOYED)) / (max(data$DAYS_EMPLOYED)-min(data$DAYS_EMPLOYED))
# # data$CNT_FAM_MEMBERS <- (data$CNT_FAM_MEMBERS-min(data$CNT_FAM_MEMBERS)) / (max(data$CNT_FAM_MEMBERS)-min(data$CNT_FAM_MEMBERS))
# 
# # ggplot(data = data, mapping = aes(x = ID, y = AMT_INCOME_TOTAL)) + geom_point()
# # heavily right skewed, but NN should not be affected by it
# #data$AMT_INCOME_TOTAL <- (data$AMT_INCOME_TOTAL-min(data$AMT_INCOME_TOTAL)) / (max(data$AMT_INCOME_TOTAL)-min(data$AMT_INCOME_TOTAL))
# 
# summary(data)
```

## Check data
```{r}
# dim(data)
# sapply(data, class)
# levels(data$status)
# 
# #Replace C and X in status
# #data$status[data$status == "C"] <- 6
# levels(data$status)[levels(data$status)=="C"] <- "6"
# #data$status[data$status == "X"] <- 7
# levels(data$status)[levels(data$status)=="X"] <- "7"
# #Convert factors into numericals
# data %<>% mutate_if(is.factor, as.numeric)
# levels(data$status)
# sapply(data, class)

# summarize the class distribution
percentage <- prop.table(table(data$status)) * 100
cbind(freq=table(data$status), percentage=percentage)

# Turn data frame into data matrix
matrix_data <- trainingSet_processed %>% select(-tail(names(trainingSet_processed), 8))
#matrix_data <- subset(data, select = c(CNT_CHILDREN, AMT_INCOME_TOTAL))
#matrix_targets <- data.matrix(trainingSet_processed[])
matrix_targets <- trainingSet_processed %>% select(tail(names(trainingSet_processed), 8))

matrix_data_test  <- testSet_processed %>% select(-tail(names(testSet_processed), 8))
matrix_targets_test  <- testSet_processed %>% select(tail(names(testSet_processed), 8))

#Subset only 100 entries for testing
#matrix_data <- matrix_data[1:100, ]
#matrix_targets <- matrix_targets[1:100, ]
```
## Build Model
```{r}
#train_data <- matrix_data
train_data <- data.matrix(matrix_data)
test_data <- data.matrix(matrix_data_test)
train_targets <- data.matrix(matrix_targets)
test_targets <- data.matrix(matrix_targets_test)

# Function to build the model
build_model <- function() {
  model <- keras_model_sequential() %>%
    #layer_batch_normalization(axis = -1L, input_shape = dim(train_data)[[2]]) %>%
    layer_dense(units = 64, activation = "relu", input_shape = dim(train_data)[[2]]) %>%
    layer_dense(units = 128, activation = "relu") %>%
    layer_dense(units = 64, activation = "relu") %>%
    layer_dense(units = 8, activation = "softmax") 

  model %>% compile(
    optimizer = optimizer_sgd(learning_rate = 0.2),
    loss = "categorical_crossentropy",
    metrics = "categorical_accuracy"
  )

}
```
## K-Fold-Validation
```{r}
# mean <- apply(matrix_data, 2, mean)
# std <- apply(matrix_data, 2, sd)
# train_data <- scale(matrix_data, center = mean, scale = std)
# test_data <- scale(matrix_data, center = mean, scale = std)
# train_targets <- matrix_targets


k <- 10
indices <- sample(1:nrow(train_data))
folds <- cut(indices, breaks = k, labels = FALSE)

num_epochs <- 1500
all_acc_histories <- NULL
for (i in 1:k) {
  cat("processing fold #", i, "\n")

  val_indices <- which(folds == i, arr.ind = TRUE)
  val_data <- train_data[val_indices,] #test_data#
  val_targets <- train_targets[val_indices,] #test_targets#
  
  partial_train_data <- train_data[-val_indices,]
  partial_train_targets <- train_targets[-val_indices,]
  model <- build_model()

  # Train the model (in silent mode, verbose=0)
  # Batch size https://stats.stackexchange.com/questions/153531/what-is-batch-size-in-neural-network
  # One epoch = one forward pass and one backward pass of all the training examples
  # Batch size = the number of training examples in one forward/backward pass. The higher the batch size, the more memory space you'll need.
  # Number of iterations = number of passes, each pass using [batch size] number of examples. To be clear, one pass = one forward pass + one backward pass (we do not count the forward pass and backward pass as two different passes).
  # Batch size 32 much faster than 1, also the smaller the batch the less accurate the estimate of the gradient will be.
  history <- model %>% fit(
    partial_train_data, partial_train_targets,
    validation_data = list(val_data, val_targets),
    epochs = num_epochs, batch_size = 512, verbose = 1
  )
  acc_history <- history$metrics$val_categorical_accuracy
  all_acc_histories <- rbind(all_acc_histories, acc_history)
}


#reticulate::py_last_error()
```

#We can then compute the average of the per-epoch ACC scores for all folds:

```{r}
average_acc_history <- data.frame(
  epoch = seq(1:ncol(all_acc_histories)),
  validation_acc = apply(all_acc_histories, 2, mean)
)

#Schreibe Ergebnis in csv, bitte den Namen abändern
#write.csv(average_acc_history, "../Doc/Versuch 3/Try 3.csv", row.names=FALSE)

max(average_acc_history$validation_acc)

library(ggplot2)
ggplot(average_acc_history, aes(x = epoch, y = validation_acc)) + geom_line()

#It may be a bit hard to see the plot due to scaling issues and relatively high variance. Let's use `geom_smooth()` to try to get a clearer picture:
ggplot(average_acc_history, aes(x = epoch, y = validation_acc)) + geom_smooth()

# Evaluate on Testset
eval <- evaluate(model, test_data, test_targets, verbose = 1)
eval
```

```{r}
# all_scores
# mean(all_scores)
```


```{r}
  ## Take Model1 and Scale data
  ##scale and preprocess training and test data
  #library(caret)
  pre_proc_val <- preProcess(train, method = c("center", "scale"))  
  train = predict(pre_proc_val, train)
  test = predict(pre_proc_val, test)
  train <- tensor_slices_dataset(train)
  test <- tensor_slices_dataset(test)
  summary(train)
  
```


```{r, results='hide'}
library(keras)
#https://www.tensorflow.org/tutorials/structured_data/feature_columns

#center and scale --> min max
#one hot encoding
# Matrix and vector

# model = tf.keras.Sequential([
#   feature_layer,
#   layers.Dense(128, activation='relu'),
#   layers.Dense(128, activation='relu'),
#   layers.Dropout(.1),
#   layers.Dense(1)
# ])
# 
# model.compile(optimizer='adam',
#               loss=tf.keras.losses.BinaryCrossentropy(from_logits=True),
#               metrics=['accuracy'])
# 
# model.fit(train_ds,
#           validation_data=val_ds,
#           epochs=10)

```



```{r}
str(train_data[[1]])
```